---
title: "Complex Sampling Design"
date: last-modified
date-format: "dddd, DD/MM/YYYY"
format: 
  html:
    theme: sandstone
    code-fold: show
    code-copy: hover
    code-overflow: wrap
    code-tools: true
    df-print: paged
    default-image-extension: svg
    embed-resources: true
    page-layout: full
    reference-location: margin
    title-block-banner: true
    title-block-style: default
    fontsize: 1.2em
    monofont: 'Fira Code'
number-sections: true
toc: true
fig-dpi: 320
dpi: 320
execute: 
  echo: fenced
---

```{r}
#| label: setup
#| echo: false

pacman::p_load(tidyverse, janitor, summarytools, gtsummary, 
               broom, rstatix, ResourceSelection, pROC, caret)
```

# Practical

## Setup Project

1. Open your RStudio

2. Create Project
  - File -> New Project...
  - New Directory -> New Project
    - set Directory Name, e.g., `Sesi 4 NHMS`
    - ensure the project directory is in default working directory, i.e., `~/Documents/RStudio`
    - click `Create Project` button

3. Copy your NHMS dataset into the working directory
  - `Files` pane -> click the `âš™ï¸ More ðŸ”»` button -> select `Show Folder in New Window`
  - Copy your NHMS dataset into the working directory
  - alternatively, you can also check you working directory, and copy the working directory into your File Explorer

```{r}
#| eval: false
#| echo: true

getwd() %>% 
  clipr::write_clip()
```


# Analysis

## Import Dataset

SPSS sav file
  - Click the spss's sav file
  - Select `Import Dataset...`
  - Copy the code into r code chunk
  - add function `as_factor(_)` 
    - Note: as_factor (from `haven` package), 
    - **NOT** as.factor (from `base` package)

```{r}

library(tidyverse)
library(haven)

nhms19ds <- read_sav("nhms19ds.sav") %>% 
  as_factor()

nhms19ds
```

Occasionally, you may have the original dataset in csv/xlsx file
  - similar step with SPSS
  - click the excel xlsx file
  - select import dataset
  - copy the code into r code chunk
  - appropriate data wrangling
    - label
    - recode
    - rename
    - relevel

```{r}
library(readxl)

xlsxnhms19 <- read_excel("NHMS_2019_CholData.xlsx")

xlsxnhms19
```

```{r}
library(haven)
library(labelled)
library(stringi)

xlsxnhms19_w <- xlsxnhms19 %>% 
  mutate(INDVID = stri_sub(indvid, 2)) %>% 
  set_value_labels(c01 = c("EJ" = -9, "TT" = -7, "N/A" = -6, 
                           "Yes" = 1, "No" = 2), 
                   c02 = c("EJ" = -9, "TT" = -7, "N/A" = -6, 
                           "Yes" = 1, "No" = 2), 
                   c03 = c("TT" = -7, "N/A" = -6, 
                           "Less than 1 year" = 1, "1 year and more" = 2), 
                   c03a = c("TT" = -7, "N/A" = -6), 
                   c04a = c("N/A" = -6, "Yes" = 1, "No" = 2), 
                   c04b = c("TT" = -7, "N/A" = -6, "Yes" = 1, "No" = 2), 
                   c04c = c("TT" = -7, "N/A" = -6, "Yes" = 1, "No" = 2), 
                   c04d = c("TT" = -7, "N/A" = -6, "Yes" = 1, "No" = 2), 
                   c05 = c("TT" = -7, "N/A" = -6, "Yes" = 1, "No" = 2), 
                   c06 = c("EJ" = -9, "TT" = -7, "N/A" = -6, 
                           "Government clinic" = 1, "Private clinic" = 2,
                           "Government hospital" = 3, "Private hospital" = 4,
                           "Pharmacy (self-medicating)" = 5, 
                           "Traditional, herbal, complementary" = 6,
                           "Didn't seek treatment" = 7),
                   u303 = c("EJ" = -9, "N/A" = -6),
                   known_chol = c("N/A" = -6, "No" = 0, "Yes" = 1),
                   undiagnosed_chol = c("N/A" = -6, "No" = 0, "Yes" = 1),
                   total_chol = c("N/A" = -6, "No" = 0, "Yes" = 1),
                   bodyweight1 = c("EJ" = -9, "TB" = -8, "N/A" = -6),
                   bodyweight2 = c("EJ" = -9, "TB" = -8, "N/A" = -6),
                   bodyheight1 = c("EJ" = -9, "TB" = -8, "N/A" = -6), 
                   bodyheight2 = c("EJ" = -9, "TB" = -8, "N/A" = -6), 
                   wc1 = c("EJ" = -9, "TB" = -8, "N/A" = -6), 
                   wc2 = c("EJ" = -9, "TB" = -8, "N/A" = -6), 
                   weight = c("N/A" = -6), 
                   height = c("N/A" = -6), 
                   wc = c("N/A" = -6)) %>% 
  set_na_values(c01 = c(-9, -7, -6), 
                c02 = c(-9, -7, -6), 
                c03 = c(-7, -6), 
                c03a = c(-7, -6), 
                c04a = -6, 
                c04b = c(-7, -6), 
                c04c = c(-7, -6), 
                c04d = c(-7, -6), 
                c05 = c(-7, -6), 
                c06 = c(-9, -7, -6), 
                u303 = c(-9, -6), 
                known_chol = -6, 
                undiagnosed_chol = -6, 
                total_chol = -6, 
                bodyweight1 = c(-9, -8, -6), 
                bodyweight2 = c(-9, -8, -6), 
                bodyheight1 = c(-9, -8, -6), 
                bodyheight2 = c(-9, -8, -6), 
                wc1 = c(-9, -8, -6), 
                wc2 = c(-9, -8, -6), 
                weight = -6, 
                height = -6, 
                wc = -6) %>% 
  set_variable_labels(state_st = "PSU",
                      ebid = "EB ID - Cluster",
                      wtfinal_ncd = "Sampling Weight",
                      c01 = "ever had total blood cholesterol level measured",
                      c02 = "ever told have high cholesterol level", 
                      c03 = "when told to have high cholesterol", 
                      c03a = "years since was told to have high cholesterol", 
                      c04a = "on medication for past 2 week", 
                      c04b = "advice for special low fat diet", 
                      c04c = "advice to loose weight", 
                      c04d = "advice to exercise", 
                      c05 = "treatment - herbal/TCM", 
                      c06 = "common place to receive treatment",
                      u303 = "Total Cholesterol (mmol/L)",
                      bodyweight1 = "Body Weight (kg)",
                      bodyweight2 = "Body Weight (kg)",
                      weight = "Body Weight (kg)",
                      bodyheight1 = "Body Height (cm)",
                      bodyheight2 = "Body Height (cm)",
                      height = "Body Height (cm)",
                      wc1 = "Waist Circumference (cm)",
                      wc2 = "Waist Circumference (cm)",
                      wc = "Waist Circumference (cm)") %>% 
  to_factor(drop_unused_labels = T, user_na_to_na = T)

xlsxnhms19_w
```

