---
title: "Statistical Tests in R: Regression"
subtitle: "Kursus R: Pengenalan dan Praktikal (Sesi 2)"
format: 
  clean-revealjs: 
    slide-number: true
    lightbox: true
    auto-stretch: false
    footer: "Linear Regression & Logistic"
author:
  - name: Mohd Azmi Bin Suliman
    orcid: 0000-0002-2125-3811
    email: azmi.suliman@moh.gov.my
    affiliations: Pusat Penyelidikan Penyakit Tak Berjangkit, Institut Kesihatan Umum
date: 2024-03-06
date-format: "dddd, DD MMMM YYYY"
embed-resources: true
execute: 
  echo: true
---

# Regression

-   Simple Linear Regression
-   Multiple Linear Regression
-   Simple Logistic Regression
-   Multiple Logistic Regression

```{r}
#| label: setup
#| echo: false

pacman::p_load(tidyverse, haven, gtsummary)
asthmads_clean <- read_sav("asthmads_clean.sav") %>% 
  as_factor()
```

```{css}
#| echo: false

code.sourceCode {
  font-size: 1.2em;
}
```

# Linear Regression

## Linear Regression

-   Use to model the relationship between a dependent (continuous) variable and one or more independent variable.
-   The formula is shown below
    -   $y$ is the dependent variable / outcome
    -   $x$ is the independent variable / predictor
    -   $\beta_0$ is the intercept (when $x$ = 0)
    -   $\beta_1$ is the slope (change in $y$ for a unit change in $x$)
    -   $\epsilon$ is the error term

$$y = \beta_0 + \beta_1x + \epsilon$$

## Linear Regression Assumptions

-   to provide reliable estimates, several key assumptions must be met
    -   Each observation/measurement were **independent**.
    -   The two variables are **continuous**.
    -   The two variables should be **approximately normally distributed**.
    -   The relationship between the two variables is **linear**.
    -   The two variables should have **equal variances**.
-   **Note**: *note the similarity between pearson correlation and linear regression*

# Practical: Linear Regression 

Finding factors associated with PEFR changes.

## Setup

-   Add a new level 1 and level 2 header
    -   `# Linear Regression`
    -   `## Simple Linear Regression`
-   Identify the variables
    -   Outcome (PEFR changes): `PEFR_Diff`

```{r}
names(asthmads_clean)
```




