---
title: "Last Session Recap"
date: last-modified
date-format: "dddd, DD/MM/YYYY"
format: 
  html:
    theme: sandstone
    code-fold: show
    code-copy: hover
    code-overflow: wrap
    code-tools: true
    df-print: paged
    default-image-extension: svg
    embed-resources: true
    page-layout: full
    reference-location: margin
    title-block-banner: true
    title-block-style: default
    fontsize: 1.2em
    monofont: 'Fira Code'
number-sections: true
toc: true
fig-dpi: 320
dpi: 320
execute: 
  echo: fenced
---

# Basic Statistical Test

In this session, several tests will be covered

1. Chi-Square Test
2. Independent t-test
3. Linear Regression
4. Logistic Regression

There will also two part, on how to conduct the tests

1. Basic, standard way
2. Using package: easier, tidier, prettier

## Chi-square Test

### Background

- test for significant relationship between two categorical variables
- test assumptions: 
  - random sampling
  - independence
  - mutually exclusive
  - <20% of cell have expected count <5
  
### Conduct Chi-square Test

Objective: To test whether employment status and gender had significant association

1. Import dataset `asthmads_clean.sav`

```{r}
library(tidyverse)
library(haven)

asthmads <- read_sav("asthmads_clean.sav") %>% 
  as_factor()
asthmads

```

2. Creata Crosstabulation

```{r}
with(asthmads, table(Gender, WorkStatus))
```

:::{.panel-tabset}

### Base too (xtabs)

```{r}
xtabs(~ Gender + WorkStatus, asthmads)

xtabs(~ Gender + WorkStatus, asthmads) %>% 
  prop.table(., margin = 2)
```

### Tidyverse & Janitor


```{r}
library(tidyverse)
library(janitor)

asthmads %>%
  count(Gender, WorkStatus)

asthmads %>%
  count(Gender, WorkStatus) %>%
  pivot_wider(names_from = WorkStatus, values_from = n, values_fill = list(n = 0))

asthmads %>%
  count(Gender, WorkStatus) %>%
  pivot_wider(names_from = WorkStatus, values_from = n, values_fill = list(n = 0)) %>%
  janitor::adorn_totals(c("row", "col"))
```


### Janitor (tabyl)

```{r}
asthmads %>%
  tabyl(Gender, WorkStatus)

asthmads %>%
  tabyl(Gender, WorkStatus) %>%
  adorn_percentages("col") %>%
  adorn_pct_formatting()
```




:::
